---
title: "Project_1_DS_6372"
author: "Jake"
date: "2023-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load Libraries, warning=FALSE, message=FALSE}
library(caret)
```
# MSDS 6372 Project 1

## Goal
We would like to predict medical costs for insurance claims.

## EDA

### Things we need to do
  * Deal with missing data if there is any
  * Clean up variable names
  * change data types
  * split data into train/validation split first (then use training data for objectives)

### EDA needs to include
  * discussion of trends between the response and predictors

```{r EDA}
# Read in data
raw_data <- read.csv("../raw_data/insurance.csv")
summary(raw_data)




```

## Objective 1

We need to create a interpretable model

  * Build a model with the main goal to identify key relationships that is highly interpret able.
  * Provide interpretation of the regression coefficients
    - Include hypotheses testing
    - Interpretation of regression coefficients
    - Confidence intervals
    - Mention the practical vs statistical significance of the predictors
  * Interpretation of at least a subset of the predictors in your final model
    - Tests on coefficients
    - assumption checking
    interpreting those coefficients
  * Feature selection
  
```{r}

```

## Objective 2

Develop a model that can predict the best and do well on future data

  * Train/validation or CV approach for model comparisons
  * Create a linear regression model WITH complexity (not just include a model with predictors that you've eliminated from objective 1)
  * Run one non-parametric model to compare to your complex model
  * Provide measures of fit for comparisons
    - MSE
    - R squared / Ajusted R squared
    - AIC & BIC
  * Use validation set to show results DO NOT tune model based on validation set
  * Feature selection and CV are a must here
  
 ```{r}
 
 
 ```
  




